---
title: "Stat 331 Portfolio"
author: "Scott Brydon"
format: 
  html: 
    theme:
      dark: darkly
    code-fold: true
    code-block-bg: '#063970'
editor: visual
---

# Table of Contents

Challenge 3 Letter B (WD-2, DVS-5, R-3)

Lab 4 Question 7 (DVS-1, DVS-2, DVS-4)

## Final Grade Reflection

## Essay Responses

## Week 6 Reflection

Now that we\'ve passed the halfway point for the class, I
feel that I\'ve become exponentially more prepared to work in a data scientist
or analyst position post-college. Therefore, I feel like my goals are being accomplished.
I feel more prepared primarily because of our recent use of complicated data
visualizations and summarization throughout these last few weeks. For example,
I feel that I\'ve gained proficiency in all the Data Visualization & Summarization
learning targets except for DVS-3. In Challenge 3, I was able to show my ability
to summarize variables across multiple groups when finding the highest
difference in music genre familiarity across white and non-white groups, which satisfies
DVS-5. In Lab 4 Question 7, I was able to successfully replicate the graph
showing the proportion of avocados by size sold in each California region. This
graph required a visualization of factor, numerical, and categorical variables
to provide enough visual information. Secondly, the x-axis labels were
difficult to read, which required the use of a dodge to stagger the labels. Lastly,
the proportion of avocados sold by size had to be calculated using mutate,
mean, and simple division. After completing the question and submitting revisions,
I was able to show proficiency in DVS-1, DVS-2, and DVS-4. In my Week 3 Reflection,
I wasn\'t familiar with the Working with Data learning target WD-5, and the Reproducibility
learning target R-3. I feel confident in my proficiency for WD-5, which I practiced
heavily throughout Week 4 with the use of semi join and anti-join. I even showed
use of it in Lab 4 Question 7. As for R-3, the summarize and select functions used
in Challenge 3 Letter B show programs that are resistant to change in inputs
due to the use of column names rather than indexing by numerical position. I
had a problem with indexing by numerical position when writing my Week 3
Reflection, so I\'m happy I\'m able to show proficiency in that learning target
now.

I\'ve demonstrated a commitment to continued learning by completing every Challenge problem and making
corrections to each Challenge I\'ve received feedback for. Throughout my solving
of these problems, I used stack overflow frequently, which has helped my Program
Efficiency. My proficiency in learning target PE-4 has drastically increased
because I\'m able to see the most efficient ways other people have been able to
utilize different functions in the tidy verse. I\'ve grown as a team member and
a leader as the labs have gotten continually harder because our group has had
to communicate more than in Weeks 1-3 to be able to complete the labs and challenge
assignments. I\'ve started to utilize discord more and have even posted a few questions
which I believe other students benefitted from as well. By doing so, I feel I\'m
helping turn discord into the helpful tool it was intended to be at the start
of this quarter. Unfortunately, I haven\'t been good about answering other people\'s
discord questions so far, which is something I need to start actively doing. I can\'t
expect an answer for every question I have if I\'m not helping provide other students
with answers as well.

I believe my current grade is an A-. I\'ve submitted full attempts/solutions for every
assignment, completed almost all assignments on time, and have demonstrated
proficiency in the vast majority of learning targets. I\'ve revised my code for
every lab or challenge I\'ve received feedback on and have shown an increase in
quality of assignments as I continue to make revisions. I\'ve been a helpful
group member in and out of class, and I ask questions when needed in class or
on discord but have been subpar about answering other student\'s questions on discord.
Lastly, I believe I\'ve given peer reviews that give constructive criticism and
enable my peers to make positive changes to their code. But I also believe I could\'ve
given better feedback in my earlier peer reviews now that I\'m not as naïve to R.

## Supporting Artifacts

```{r}
#| message = FALSE

library(tidyverse)
library(here)
```

1.  Challenge 3 Letter B (WD-2, DVS-5, R-3)

```{r}
#| message = FALSE

hiphop <- read_csv(here("Week 3", "Lab 3", "hiphop.csv"))

new_hiphop <- hiphop |> 
  mutate(
    white_or_nonwhite = 
      if_else(ethnic != "white", "non-white", "white"),
    white_or_nonwhite = 
      as.factor(white_or_nonwhite)
  )

new_hiphop |> 
  select(intl, vocal, classical, 
         folk, rock, country, 
         pop, alternative, hiphop, 
         unclassifiable, white_or_nonwhite) |> 
  group_by(white_or_nonwhite) |> 
  summarise(across(intl:unclassifiable, mean)) |> 
  summarise(across(intl:unclassifiable, diff))
```

2.  Lab 4 Question 7 (WD-5, R-3, DVS-1, DVS-2, DVS-4,)

```{r}
#| message = FALSE

avocados <- read_csv(here("Week 4", "Lab 4", "avocado.csv"))

california_regions <- tibble(CaliforniaRegions = c(
  "SanFrancisco", "SanDiego", "LosAngeles","Sacramento")
  )

avocados <- avocados |> 
  rename("Small" = "4046",
         "Large" = "4225",
         "Extra Large" = "4770")

avocados_longer <- avocados |> 
  semi_join(california_regions, by = c("region" = "CaliforniaRegions")) |> 
  pivot_longer("Small":"Extra Large",
               names_to = "Avocado Size",
               values_to = "Number Avocados") |> 
  mutate(Proportion = `Number Avocados` / mean(`Total Volume`))
  

avocados_longer |> 
  ggplot(mapping = aes(x = region, y = Proportion, fill = `Avocado Size`)) +
  geom_bar(position = position_fill(reverse = TRUE), stat = "identity") +
  labs(x = "Regions of CA", y = "Proportion of Mean Avocados Sold") +
  #Found function below on Stack Overflow 
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +
  facet_wrap(~ type)
```

### Demonstration of Learning Targets

### Demonstration of Continued Learning

#### Revising My Thinking

#### Extending My Thinking

### Growth as a Team Member
