---
title: "Lab 9"
author: "Scott Brydon"
format: 
  html: 
    theme:
      dark: darkly
    code-fold: true
    code-block-bg: '#063970'
editor: visual
---

```{r}
#| message: false
library(tidyverse)
library(here)
StateNames_A <- read_csv(here("Week 9", "Lab 9", "StateNames_A.csv"))
```

#### Summarizing and Visualizing the Number of Allisons

1\.

```{r}
StateNames_A <- StateNames_A |> 
  rename("Sex at Birth" = "Gender") 

StateNames_A |> 
  pivot_wider(
    names_from = `Sex at Birth`,
    values_from = Count
  ) |> 
  filter(Name == "Allison") |> 
  group_by(State) |> 
  summarize(across(
    F:M, sum)
    ) |>  
  mutate(
    across(F:M, ~replace_na(., 0))
    )

```

2/3.

```{r}
Sum_Allisons <- StateNames_A |> 
  filter(`Sex at Birth` == "F",
         Name == "Allison") |> 
  group_by(Year) |> 
  summarize(Total_Allison = sum(Count))

Sum_Allisons |> 
  ggplot(mapping = aes(x = Year, y = Total_Allison)) +
  geom_col() +
  labs(y = "", 
       title = "Number of Females named Allison per Year in the US")
```

#### Modeling the Number of Allison's

4/5.

```{r}
Allison_Reg <- Sum_Allisons %>%
  lm(Total_Allison ~ Year, data = .)

Allison_Reg |> 
  broom::tidy()

Allison_Reg |> 
  broom::glance()

Sum_Allisons |> 
  ggplot(mapping = aes(x = Year, y = Total_Allison)) + 
  geom_point() +
  stat_smooth(method = "lm")
```

6.  ŷ = 3881.886 - 1.876(Year)
7.  

```{r}
Allison_Reg |> 
  broom::augment() |> 
  ggplot(mapping = aes(x = .fitted, y = .resid)) +
  geom_point()
```

The majority of the residuals look to be random around zero, which implies the model is probably a linear relationship. There does look to be a slight negative linear pattern in the residuals due to the outliers in the top left region.

8\.

We conclude from this model that the total number of females at birth named Allison decreases by an average of -101.52 each additional year in each individual state. The intercept is meaningless since it represents the total number of females named Allison at birth in a state in year zero.

#### Spelling by State

1\.

```{r}
StateNames_A |>
  filter(`Sex at Birth` == "M", 
         Name %in% c("Allan", "Alan", "Allen")
         ) |> 
  ggplot(mapping = aes(x = Year, y = Count, colour = Name)) +
  geom_smooth() +
  labs(title = "Popularity of different Allan name spellings over time")
```

2\.

```{r}
StateNames_A |> 
  filter(Name %in% c("Allan", "Alan", "Allen")) |> 
  pivot_wider(
    names_from = Name, 
    values_from = Count
    ) |> 
  filter(`Sex at Birth` == "M",
         Year == 2000,
         State %in% c("PA", "CA")) |> 
  mutate(
    across(Allan:Allen, ~replace_na(., 0)),
    AllenPercent = round(Allen/ rowSums(across(Allen:Allan)), 2),
    AlanPercent = round(Alan/ rowSums(across(Allen:Allan)), 2),
    AllanPercent = round(Allan/ rowSums(across(Allen:Allan)), 2)
    )
```
