---
title: "Lab 7"
author: "Scott Brydon"
format: 
  html: 
    self-contained: true
    theme:
      dark: darkly
    code-fold: true
    code-block-bg: '#063970'
editor: visual
---

```{r}
#| message: false
library(tidyverse)
library(here)
BlackfootFish <- read_csv(here("data", "BlackfootFish.csv"))
```

#### Task 1

```{r}
Count_NAs <- function(data) {
  sum(is.na(data))
}

BlackfootFish |> 
  summarize(across(.cols = everything(),
         .fns = Count_NAs))
```

#### Task 2

```{r}
#| message: false

BlackfootFish |> 
  group_by(year, trip, section) |> 
  summarize(across(.cols = everything(),
         .fns = Count_NAs)) |> 
  ggplot(mapping = aes(x = year, y = weight, fill = section)) +
  geom_col() +
  coord_flip() +
  facet_wrap(~ trip)
  
```

## Part 2:

#### Task 1/2/3

```{r}
rescale_01 <- function(vector){
  stopifnot(is.numeric(vector), length(vector) > 1)
  
  min_vec <- min(vector, na.rm = TRUE)
  max_vec <- max(vector, na.rm = TRUE)
  vector <-  (vector - min_vec) / (max_vec - min_vec)
  
  return(vector)
  
}

x <- c(1:25, NA)
rescale_01(x)
```

#### Task 4

```{r}
#| warning: false
#| layout-ncol: 2

BlackfootFish |> 
  ggplot(mapping = aes(
    x = weight, 
    y = length,
    color = species)
    ) +
  geom_boxplot()

BlackfootFish |> 
  ggplot(mapping = aes(
    x = weight, 
    y = rescale_01(length),
    color = species
    )) +
  geom_boxplot() +
  labs(y = "Rescaled Length")
```

#### Task 5/6

```{r}
rescale_column <- function(data, vectors) {
  stopifnot(is.data.frame(data))
  
  data %>%
    mutate( 
      across(
      .cols = {{ vectors }}, 
      .fns = rescale_01
      )
    )
}

rescale_column(BlackfootFish, length:weight)
```

**Lab 7 Revisions**

**Task 2:** When I had originally formatted the ggplot for the number of NAs in the weight column, I was struggling with the correct plot to use because I had been considering the year as a purely numerical variable in this case, and there tried using geom_line and geom_point. Once I realized that year was acting more as a categorical variable because there would never be any value at a fraction of a year, I reverted to using a bar plot. Once I got to this point, I tried switching the year to the y-axis in the aesthetics argument, but it didn't produce the horizontal bar graph desired. I assumed this was the best that could be done using ggplot because I hadn't remembered the usefulness of the coord_flip function. Until you hinted at the possibility of years being on the y-axis in my corrections, I never would have realized I could have flipped it.

**Task 1/2/3:** For the stopifnot function, I didn't understand the argument structure for the function. I assumed it was a function that reduced the need to write convoluted if statement. Therefore, I figured it would follow the same logical operator structure and require an or symbol between the arguments being tested.

The redundancy in the rescale function was a result of me assuming that since the minimum function is already defined, that I couldn't replace it with a simpler argument. I had been using the thought process of developing functions to minimize redundancy and knew I couldn't make another minimize function. This now seems obvious, but I hadn't thought at the time to just define the minimum and maximum as objects precalculated in the function. This happens a lot to me in this class, where I'll get so homed in on a specific learning target or way of doing things that I'll overcomplicated my thinking.

**Task 4:** Whenever I think of the simplest type of graph, I gravitate towards the scatterplot. This can be counterintuitive because often the purpose of the visual and the interpretation don't require a scatterplot. That happens to be the case here, where a scatterplot makes the data almost overlap and extremely hard to interpret. If there were slight changes to the data when I rescaled it, it would've been hard to see small movements in the points too. I decided to change to a boxplot because you commented we only want to see the distributions of weight, which would be shown by the quantiles of a boxplot. In general, I need to become better at answering the question in a straightforward manner.

***Feedback Provided:***

*In this class, we use a style guide which recommends the use of an assignment arrow over an `=` sign for object creation.*

*Careful! The code chunk options are specified after a `:` and uses **lowercase** letters -- `false` and `true`. Using capitals or a mix will cause rendering errors. I was unable to render your document due to your option (`#| message = FALSE`) not using `:` and lowercase letters.*

*Wouldn\'t it be nice if there were continuous values for the years? :) What is being plotted on the y-axis?*

*A `,` in `stopifnot()` is the same as an `|`. So, I don\'t see a reason not to use the default argument structure for the function.*

*Great function! I\'d press you to think about the redundancy in your function. Is the same function being called multiple times? If so, how can you remove this redundancy?*

*The only variable I am concerned with is `weight`. I want visualizations that make it as simple as possible to compare the distributions of `weight` before and after scaling!*
