---
title: "Challenge 3"
author: "Scott Brydon"
format: 
  html: 
    theme:
      dark: darkly
    code-fold: true
    #Example of coloring code blocks
    code-block-bg: '#063970'
editor: visual
execute:
  echo: true
  error: true
---

#### Published Comparisons

```{r message = FALSE}
library(tidyverse)
library(here)
hiphop <- read_csv(here("Week 3", "Lab 3", "hiphop.csv"))
```

```{r}
new_hiphop <- hiphop |> 
  mutate(
    white_or_nonwhite = 
      if_else(ethnic != "white", "non-white", "white"),
    white_or_nonwhite = 
      as.factor(white_or_nonwhite)
  )
```

```{r}
#a. Male versus Female
new_hiphop |> 
  select(intl, vocal, classical, 
         folk, rock, country, 
         pop, alternative, hiphop, 
         unclassifiable, sex) |> 
  group_by(sex) |> 
  summarise(across(intl:unclassifiable, mean)) |> 
  summarise(across(intl:unclassifiable, diff))

#b. White versus Non-White
new_hiphop |> 
  select(intl, vocal, classical, 
         folk, rock, country, 
         pop, alternative, hiphop, 
         unclassifiable, white_or_nonwhite) |> 
  group_by(white_or_nonwhite) |> 
  summarise(across(intl:unclassifiable, mean)) |> 
  summarise(across(intl:unclassifiable, diff))
```

A. For Males vs Females, the rock genre differentiates the groups the most with a difference of roughly 1.1.

B. For Whites vs Non-Whites, the international music genre differentiates the groups the most with a difference of roughly 1.6.

#### Study Design Critique

My first major concern is the use of very stereotypical pop-culture variables to represent one's familiarity with "black pop-culture". It seems very incorrect to describe black pop-culture as simply one's knowledge of jay-z, boondocks, Charles Barkley, etc. Additionally, coining this term African-American English is stereotyping such a large portion of African-Americans. By making this dialect apply to all African-Americans via the name of it, the study makes the assumption all African-Americans talk this way.
